Boosting Bayesian Parameter Inference of Nonlinear Stochastic Differential Equation Models by Hamilton Equation Scale Separation

### 2015-10-22 TeXFiles/Manuscript_Albert_Ulzega_Stoop.tex
minor changes in tex file

Modeling a dynamical process starts with a basic model of the process, which usually is obtained from more or less deep insight into the nature of the process. The next step then is the determination of the parameters of the model, based on observed data, which is generally a highly nontrivial task, in particular when complex behavior of such systems needs to be predicted, or when the measurements are noisy. 
A minimal example is a perceptron [XXX], the basic element of a neuronal network, that predicts the double cosine value associated with the input of the corresponding sine function plus the sine's value at a fixed earlier time. While this task can easily be achieved for clean data using gradient descent learning, for noisy input data, this is largely impossible, as noise cannot be learned. The result is a distribution of potential parameter values (Fig.~(\ref{neural})).
%\begin{figure}[h]
%    \centering
%    \includegraphics[width=1.0\textwidth]{neuron.pdf}
%    \caption{General problem setting: Parameter estimation (synaptic weights $w_1$,$w_2$) of a perceptron, from a) noiseless, b) noisy data (noise sampled from a flat distribution over the  interval $[-0.2,0.2]$). Whereas for noiseless data the estimates perfectly converge, for noisy data the estimates  the system attempts to also include the noise, leading to a nontrivial distribution of the parameter estimates, and rendering the extraction of the optimal parameters a nontrivial task. Red circles: location of the optimal parameters in the noiseless case.}
%    \label{neural}
%\end{figure}
%If the physics underlying the evolution of the system follows a differential equation,
%the search for the optimal parameters and their typical distribution is called Bayesian inference of a stochastic differential equation (SDE).

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{FigPolymer2.pdf}
    \caption{Simulated polymer chain dynamics, with $n+1=11$ data points (large circles) and $j-1=9$ intermediate beads (small circles). For other parameters see Section~\ref{numerical_results}. Bottom: initial state, where intermediate beads are on a linear interpolation between data points. Top: polymer after $T=1000$ iterations of the propagation algorithm (dotted line: initial configuration). Clearly the new configuration is mostly determined by the dynamics of the light-mass intermediate beads, while the heavy-mass data points move to a much lesser extent.}
    \label{fig:polymer}
\end{figure}

>>>>

Modeling a dynamical process starts with a basic model of the process, which usually is obtained from more or less deep insight into the nature of the process. The next step then is the determination of the parameters of the model, based on observed data, which is generally a highly nontrivial task, in particular when complex behavior of such systems needs to be predicted, or when the measurements are noisy.
A minimal example is a perceptron [XXX], the basic element of a neuronal network, that predicts the double cosine value associated with the input of the corresponding sine function plus the sine's value at a fixed earlier time. While this task can easily be achieved for clean data using gradient descent learning, for noisy input data, this is largely impossible, as noise cannot be learned. The result is a distribution of potential parameter values (Fig.~(\ref{neural})).
\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\textwidth]{neuron.pdf}
    \caption{General problem setting: Parameter estimation (synaptic weights $w_1$,$w_2$) of a perceptron, from a) noiseless, b) noisy data (noise sampled from a flat distribution over the  interval $[-0.2,0.2]$). Whereas for noiseless data the estimates perfectly converge, for noisy data the estimates  the system attempts to also include the noise, leading to a nontrivial distribution of the parameter estimates, and rendering the extraction of the optimal parameters a nontrivial task. Red circles: location of the optimal parameters in the noiseless case.}
    \label{neural}
\end{figure}
If the physics underlying the evolution of the system follows a differential equation,
the search for the optimal parameters and their typical distribution is called Bayesian inference of a stochastic differential equation (SDE).

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{FigPolymer2.pdf}
    \caption{Simulated polymer chain dynamics, with $n+1=11$ data points (large circles) and $j-1=9$ intermediate beads (small circles). For other parameters see Section~\ref{numerical_results}. Bottom: initial state, where intermediate beads are on a linear interpolation between data points. Top: polymer after $N_{MC}=1000$ iterations of the propagation algorithm (dotted line: initial configuration). Clearly the new configuration is mostly determined by the dynamics of the light-mass intermediate beads, while the heavy-mass data points move to a much lesser extent.}
    \label{fig:polymer}
\end{figure}
